---
title: 'Reproducible Data: Week 2 Assignment'
output: 
  html_document:
    keep_md: true
---

Loading and preprocessing the data

```{r}
setwd('C:/Users/Allen/Google Drive/Coursera/Reproducible Data/Week 2/Assignment/data')
dat<-read.csv('activity.csv')
dat_original<-read.csv('activity.csv')
dat<-na.omit(dat)
```


**Question 1:**
What is mean total number of steps taken per day?

```{r}
totalStepsPerDay<-aggregate(dat$steps,by=list(dat$date),FUN='sum')

meanStepsPerDay<-aggregate(dat$steps,by=list(dat$date),FUN='mean')

q1_dataset<-merge(totalStepsPerDay,meanStepsPerDay,by.x='Group.1', by.y='Group.1')

medianStepsPerDay<-aggregate(dat$steps,by=list(dat$date),FUN='median')

q1_dataset<-merge(q1_dataset,medianStepsPerDay,by.x='Group.1', by.y='Group.1')

colnames(q1_dataset)<-c('date','total','mean','median')
```


Histogram of Total Steps Per Day

```{r}
hist(q1_dataset$total)
```


Total, Mean, and Median by Date Report

```{r}
print(q1_dataset)
```


**Question 2:**
Graph of AVG Daily Activity Pattern

```{r}
library(ggplot2)
q2_dataset<-aggregate(dat$steps,by=list(dat$interval),mean)
colnames(q2_dataset)<-c('Interval','AvgSteps')
ggplot(q2_dataset, aes(Interval, AvgSteps))+geom_line()
```

Interval with the Highest AVG Steps

```{r}
q2_dataset[q2_dataset$AvgSteps==max(q2_dataset$AvgSteps),]
```

**Question 3: **
Imputing missing values

Number of NA's

```{r}
sum(is.na(dat_original$steps))
```

Use overall mean as the NAs

```{r}
dat_original$steps[is.na(dat_original$steps)] <- mean(dat$steps)

totalStepsPerDay_q3<-aggregate(dat_original$steps,by=list(dat_original$date),FUN='sum')

meanStepsPerDay_q3<-aggregate(dat_original$steps,by=list(dat_original$date),FUN='mean')

q3_dataset<-merge(totalStepsPerDay_q3,meanStepsPerDay_q3,by.x='Group.1', by.y='Group.1')

medianStepsPerDay_q3<-aggregate(dat_original$steps,by=list(dat_original$date),FUN='median')

q3_dataset<-merge(q3_dataset,medianStepsPerDay_q3,by.x='Group.1', by.y='Group.1')

colnames(q3_dataset)<-c('date','total','mean','median')

```

Histogram of Total Steps Per Day

```{r}
hist(q3_dataset$total)
```


Total, Mean, and Median by Date Report After Imputing Missing Values

```{r}
q3_dataset_final<-merge(q3_dataset,q1_dataset,by.x='date',by.y='date',all.x=TRUE)
colnames(q3_dataset_final)<-c('date','tot_incl_na','mean_incl_na','med_incl_na','tot_no_na','mean_no_na','med_no_na')
q3_dataset_final$tot_incl_na<-round(q3_dataset_final$tot_incl_na,3)
q3_dataset_final$mean_incl_na<-round(q3_dataset_final$mean_incl_na,3)
q3_dataset_final$med_incl_na<-round(q3_dataset_final$med_incl_na,3)
q3_dataset_final$tot_no_na<-round(q3_dataset_final$tot_no_na,3)
q3_dataset_final$mean_no_na<-round(q3_dataset_final$mean_no_na,3)
q3_dataset_final$med_no_na<-round(q3_dataset_final$med_no_na,3)
q3_dataset_final$totDifference<-q3_dataset_final$tot_incl_na-q3_dataset_final$tot_no_na
print(q3_dataset_final)

```

**Question 4:**
Weekday vs Weekend

```{r}
q4_dataset<-dat_original
q4_dataset$weekType<-sapply(q4_dataset$date, function(x) if(weekdays(as.Date(x))=='Saturday'|weekdays(as.Date(x))=='Sunday'){'Weekend'
  }else {
      'Weekday'
    })

q4_dataset_agg<-aggregate(steps~interval+weekType,data=q4_dataset,mean)

ggplot(q4_dataset_agg, aes(interval, steps))+geom_line()+ facet_grid(weekType ~ .)
```